---
title: "Fish Depth Preliminary Results"
author: "Hannah L. Owens"
date: "12/27/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "~/Dropbox/DeepFishOccurrences/")
```

## Fitting data for use

First, I read in the data and clean the maximum and minimum depth columns, getting rid of units and converting everything (that's noted) into meters. It is assumed that measurements without units are in meters.

```{r read in data and process max and min depth values}
library(stringr)

fishData <- read.csv("data/AA_MASTER_Fishes_Global_200-8000m.csv")

# There are 187 columns in this nonsense. I'm getting rid of some.
fishData <- fishData[, c("dwc.family", "dwc.genus", "dwc.scientificName",
                         "dwc.verbatimDepth", "dwc.maximumDepthInMeters", "dwc.minimumDepthInMeters")]

# Cleaning scientific names
fishData$dwc.scientificName <- word(fishData$dwc.scientificName, 1,2, sep=" ")

# Fix blanks
fishData[fishData==""] <- NA

# Fix max and min columns
convertFMtoM <- function(input){
  output <- str_split(input, pattern = "fms")[[1]][1]
  output <- as.numeric(output) * 1.8288
  return(output)
}

convertFTtoM <- function(input){
  output <- str_split(input, pattern = "ft")[[1]][1]
  output <- as.numeric(output) * 0.3048
  return(output)
}

for (i in seq(1, nrow(fishData))){
  max <- fishData$dwc.maximumDepthInMeters[i]
  min <- fishData$dwc.minimumDepthInMeters[i]
  # Get rid of units
  if(is.na(max)){
  } else {
    if(str_detect(max, "-")){
    fishData$dwc.verbatimDepth[i] <- max
    max <- NA
    } else {
      # Get rid of random trash
      if(str_detect(max, ">")){
        max <- gsub(pattern = ">\\s*", replace = "", max)
      }
      if(str_detect(max, "ca")){
        max <- gsub(pattern = "ca\\s*", replace = "", max)
      } 
      if(str_detect(max, "c.")){
        max <- gsub(pattern = "c.\\s*", replace = "", max)
      }
      if(str_detect(max, "approx")){
        max <- gsub(pattern = "a\\w*\\s*", replace = "", max)
      }
      if(str_detect(max, "~")){
        max <- gsub(pattern = "~\\s*", replace = "", max)
      }
      if(str_detect(max, ",")){
        max <- gsub(pattern = ",", replace = "", max)
      }
      # Fix units
      if(any(str_detect(max, pattern = c("feet", "ft", "'")))){
        max <- gsub(pattern = "\\s*f\\w*", replace = "ft", max)
        max <- gsub(pattern = "\\s*'*", replace = "ft", max)
        max <- convertFTtoM(max)
      } else if(any(str_detect(string = max, pattern = c("fm", "fms", "fathoms")))){
        max <- gsub(pattern = "\\s*f\\w*", replace = "fms", max)
        max <- convertFMtoM(max)
      } else if(any(str_detect(max, pattern = c("m", "meters", "M")))){
        max <- gsub(pattern = "\\s*m\\w*", replace = "m", max)
        max <- gsub(pattern = "\\s*M\\w*", replace = "m", max)
        max <- str_split(max, "m", simplify = T)[1,1]
        tryCatch({max <- as.numeric(max)}, warning = function(w) {cat("Index:", i, "max:", max, "\n")})
      } else {
        # For everyone else
        tryCatch({max <- as.numeric(max)}, warning = function(w) {cat("Index:", i, "max:", max, "\n")})
      }
    }
  }
  
  if(is.na(min)){
  } else {
    if(str_detect(min, "-")){
      fishData$dwc.verbatimDepth[i] <- min
      min <- NA
    } else {
    # Get rid of random trash
      if(str_detect(min, ">")){
        min <- gsub(pattern = ">\\s*", replace = "", min)
      }
      if(str_detect(min, "ca")){
        min <- gsub(pattern = "ca\\s*", replace = "", min)
      } 
      if(str_detect(min, "c.")){
        min <- gsub(pattern = "c.\\s*", replace = "", min)
      }
      if(str_detect(min, "approx")){
        min <- gsub(pattern = "a\\w*\\s*", replace = "", min)
      }
      if(str_detect(min, "~")){
        min <- gsub(pattern = "~\\s*", replace = "", min)
      }
      if(str_detect(min, ",")){
        min <- gsub(pattern = ",", replace = "", min)
      }
      # Fix units
      if(any(str_detect(min, pattern = c("feet", "ft", "'")))){
        min <- gsub(pattern = "\\s*f\\w*", replace = "ft", min)
        min <- gsub(pattern = "\\s*'*", replace = "ft", min)
        min <- convertFTtoM(min)
      } else if(any(str_detect(string = min, pattern = c("fm", "fms", "fathoms")))){
        min <- gsub(pattern = "\\s*f\\w*", replace = "fms", min)
        min <- convertFMtoM(min)
      } else if(any(str_detect(min, pattern = c("m", "meters", "M")))){
        min <- gsub(pattern = "\\s*m\\w*", replace = "m", min)
        min <- gsub(pattern = "\\s*M\\w*", replace = "m", min)
        min <- str_split(min, "m", simplify = T)[1,1]
        tryCatch({min <- as.numeric(min)}, warning = function(w) {cat("Index:", i, "min:", min, "\n")})
      } else {
      # For everyone else
        tryCatch({min <- as.numeric(min)}, warning = function(w) {cat("Index:", i, "min:", min, "\n")})
      }
    }
  }
  
  # Check for max/min inversion and assign to rightful place in table
  if(all(!is.na(c(max,min)))){
    fishData$dwc.maximumDepthInMeters[i] <- max(c(max,min))
    fishData$dwc.minimumDepthInMeters[i] <- min(c(max,min))
  } else{
    fishData$dwc.maximumDepthInMeters[i] <- max
    fishData$dwc.minimumDepthInMeters[i] <- min
  }
}
```

Next, for species without maximum and/or minimum depths, we try to extract additional information from the verbatim depth data. Assumptions again include a lack of units implying the units are in meters.

``` {r process verbatim depths}
# Mining verbatim depth for precious missing information to add to max and min columns
for (i in seq(1, nrow(fishData))){
  entry <- fishData$dwc.verbatimDepth[i]
  entry <- gsub(pattern = "\\.$", replacement = "", entry)
  max <- fishData$dwc.maximumDepthInMeters[i]
  min <- fishData$dwc.minimumDepthInMeters[i]
  if(all(any(is.na(c(max, min))), !is.na(entry))){
    if(str_detect(entry, pattern = "\\d+")){
      if(str_detect(entry, pattern = "minute", negate = T)){
        # Pull out units
        if(str_detect(entry, pattern = "\\d+\\s*[fmM]+")){
          units <- str_extract(entry, "[fmM]\\w*+$")
          if(any(str_detect(units, pattern = c("fa", "fathoms", "fm")))){
            units <- "fm"
          } else if(any(str_detect(units, pattern = c("m", "M")))){
            units <- "m"
          } 
        } else {
          units <- "m"
        }
        # Pull out max and min
        if(str_detect(entry, pattern = "-")){
          entry <- gsub(entry, pattern = "\\s*\\D+$", replacement = "")
          spl <- str_split(entry, pattern = "-", simplify = T)[1:2]
          max <- as.numeric(max(c(max, spl)))
          min <- as.numeric(min(c(min, spl)))
        } else if (str_detect(entry, pattern = "\\d\\s\\d")){
          entry <- gsub(entry, pattern = "\\s*\\D+$", replacement = "")
          entry <- gsub(entry, pattern = ",", replacement = "")
          spl <- str_split(entry, pattern = "\\s", simplify = T)[1:2]
          max <- as.numeric(max(c(max, spl)))
          min <- as.numeric(min(c(min, spl)))
        } else{
          entry <- gsub(entry, pattern = ",", replacement = "")
          entry <- str_match(entry, "(\\d+\\.*\\d*)\\s*\\w*$")[2]
          max <- as.numeric(max(c(max, entry), na.rm = T))
          min <- as.numeric(min(c(min, entry), na.rm = T))
        }
        # Convert max and min
        if(!is.na(units)){
          if(units == "ft"){
            if(!is.na(min)){
              min <- convertFTtoM(min)
            }
            if(!is.na(max)){
              max <- convertFTtoM(max)
            }
          } else if(units == "fm"){
            if(!is.na(min)){
              min <- convertFMtoM(min)
            }
            if(!is.na(max)){
              max <- convertFMtoM(max)
            }
          }
        }
      }
    }
  }
  fishData$dwc.maximumDepthInMeters[i] <- max
  fishData$dwc.minimumDepthInMeters[i] <- min
  rm(min, max)
}
```
Finally, I calculate a column with a single depth measurement, the median between minimum and maximum depth. In cases where only the minimum or maximum are given, this number is used. I then save the .csv file, because dang, that's a lot of records to process.

```{r calculate median depth column}

```

## Depth by family

```{r depth by family}
ggplot(plot.data, aes(x=group, y=Oxygen)) + 
  geom_violin(fill=cols[2], ) +
  theme_classic(base_size = 40) +
  theme(axis.title = element_blank(), 
        text = element_text(family = "Optima"), 
        axis.text = element_text(size = rel(1.1))) +
  ggtitle("Dissolved Oxygen (µmol/kg)") + 
  geom_boxplot(width=0.1)
```

## Depth by genus


```{r pressure}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
